### 2008-10-08 trunk/Papers/rmrs-08/rmrs.tex
picture fixes and font change
quick changes to Extensions

\begin{figure}
  \centering
  \includegraphics[width=6cm]{pic-more-stuff}
  \caption{Another tree which satisfies (\ref{ex:cat-pos}).}
  \label{fig:fat-black-cat}
\end{figure}

@@

In the earlier \rmrs\ literature, each term of an \rmrs\ is equipped
with a \emph{sort}.  For instance, individual variables are typically
assigned the sort $x$, event variables the sort $e$, and holes the
sort $h$; sometimes subsorts of these are considered (e.g.\ $e_{PAST}$
for events in the past tense), and all the sorts arranged in a sort
hierarchy ${\cal S}$.

While we have defined \rmrs\ without sorts to simplify the
presentation, it is straightforward to add sorts to the definition.
For this, one assumes that the signature $\Sigma$ is sorted, i.e.\
assigns a sort $s_1\times\ldots s_n\rightarrow s$ to each constructor,
where $n$ is the constructor's arity (possibly zero) and
$s,s_1,\ldots,s_n\in {\cal S}$ are atomic sorts.  We can then
restrict the models of \rmrs\ to trees that are \emph{well-sorted} in
the usual sense, i.e.\ those in which we can infer a sort for each
subtree, and require that the variable assignment functions likewise
respect the sorts.  If we then modify Def.~\ref{defn:solved-forms}
such that the constructor $p$ of sufficiently high arity is also
consistent with the sorts of the known arguments---i.e., if $p$ has
sort $s_1 \times \ldots \times s_n \rightarrow s$ and the \rmrs\
contains an atom $\ARG_{\{k\}}(Y,i)$ and $i$ is of sort $s'$, then
$s'$ is a subsort of $s_k$---all the above propositions about solved
forms remain true.

Another extension is to replace
the qeq atom $X \qeqdom Y$ with dominance $X \dom Y$, 
which is satisfied iff there is a path
from $\alpha(X)$ to $\alpha(Y)$ in the tree.  Dominance is clearly a
less restrictive condition than qeq, but it has been shown that the
distinction between the two relations is irrelevant in \mrs s that were
derived from a corpus by a deep grammar \cite{FucKolNieTha04}.  In our
above definition of \rmrs, it is possible to replace qeq by dominance;
this changes \rmrs\ to a different logic (let's call it
Dominance-\rmrs), for which however all the above propositions about
solved forms still hold if ``qeq'' is systematically replaced by
``dominance''.

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{pic-more-stuff}
  \caption{Another tree which satisfies (\ref{ex:cat-pos}).}
  \label{fig:fat-black-cat}
\end{figure}

@@

So far we have included the dominance relation from
\cite{egg:etal:2001} in the syntax and semantics of the formalism and
not the qeq relation from \cite{copestake:etal:2005}.  This is partly
because dominance is more expressive than qeq; for instance if a
dependency parser links a determiner to a noun which in turn is
linked to a verb, then the dominance relation is required to express
the fact that
that the argument to the predicate that's introduced by the verb is bound by
the quantifier introduced by 
the determiner (see earlier discussion).  However, it is very
straightforward to add qeq to the syntax of the language and to extend
the model theory accordingly.  This extension adds a new atom $X=_q Y$
to Def.~\ref{defn:rmrs-syntax}, and $\tau,\alpha,g,\sigma$ will
satisfy $X=_q Y$ iff $\alpha(X)$ and $\alpha(Y)$ are separated by
zero or more nodes in $\tau$, each of which has a third child that is a
quantifier.  It is possible to replace dominance by qeq in the above
definition of \rmrs\ as well; and all the above propositions about
solved forms still hold if ``dominance'' is systematically replaced
with ``qeq''.

Furthermore, grammar developers such as those in the {\sc delphin}
community typically adopt conventions that restrict them to a
fragment of the language from Def.~\ref{defn:rmrs-syntax} (once
qeq is added to it), or they restrict attention to only a subset of
the models (e.g., any variable that equals the variable
that's bound by a quantifier is also bound by that quantifier). 
Our formalism provides a general framework
into which all these various fragments fit, and it's a matter of
future work to investigate how these fragments afford efficient
algorithms for performing inference over the \rmrs s.

Another feature of the existing \rmrs\ literature is that each term of
an \rmrs\ is equipped 
with a \emph{sort}.  For instance, individual variables are typically
assigned the sort $x$, event variables the sort $e$, and holes the
sort $h$; sometimes subsorts of these are considered (e.g.\ $e_{PAST}$
for events in the past tense), and all the sorts arranged in a sort
hierarchy ${\cal S}$.
While we have defined \rmrs\ without sorts to simplify the
presentation, it is straightforward to add sorts to the definition.
For this, one assumes that the signature $\Sigma$ is sorted, i.e.\
assigns a sort $s_1\times\ldots s_n\rightarrow s$ to each constructor,
where $n$ is the constructor's arity (possibly zero) and
$s,s_1,\ldots,s_n\in {\cal S}$ are atomic sorts.  We can then
restrict the models of \rmrs\ to trees that are \emph{well-sorted} in
the usual sense, i.e.\ those in which we can infer a sort for each
subtree, and require that the variable assignment functions likewise
respect the sorts.  If we then modify Def.~\ref{defn:solved-forms}
such that the constructor $p$ of sufficiently high arity is also
consistent with the sorts of the known arguments---i.e., if $p$ has
sort $s_1 \times \ldots \times s_n \rightarrow s$ and the \rmrs\
contains an atom $\ARG_{\{k\}}(Y,i)$ and $i$ is of sort $s'$, then
$s'$ is a subsort of $s_k$---all the above propositions about solved
forms remain true.
